---
title: "GAMM"
author: "Mark Jones"
date: "`r Sys.time()`"
output:
  html_document:
    classoption: landscape
    css: style.css
    number_sections: yes
    self_contained: yes
    theme: united
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document:
    toc: yes
    toc_depth: '3'
geometry: left=0.2cm,right=0.2cm,top=1cm,bottom=1cm
editor_options:
  chunk_output_type: console
classoption: landscape
---



```{r}
library(mgcv)
library(nlme)
## load custom functions
source("derivFun.R")
source("tsDiagGamm.R")


URL <- url("http://www.cru.uea.ac.uk/cru/data/temperature/HadCRUT3v-gl.dat")
gtemp <- read.table(URL, fill = TRUE)
## Don't need the even rows
gtemp <- gtemp[-seq(2, nrow(gtemp), by = 2), ]
## set the Year as rownames
rownames(gtemp) <- gtemp[,1]
## Add colnames
colnames(gtemp) <- c("Year", month.abb, "Annual")
## Data for 2011 incomplete so work only with 1850-2010 data series
gtemp <- gtemp[-nrow(gtemp), ]
## Plot the data
ylab <- expression(Temperature~Anomaly~(1961-1990)~degree*C)
plot(Annual ~ Year, data = gtemp, type = "o", ylab = ylab)
```






```{r}
m1 <- gamm(Annual ~ s(Year, k = 20), data = gtemp)
summary(m1$gam)
acf(resid(m1$lme, type = "normalized"))
## ...wait... look at the plot, only then do...
pacf(resid(m1$lme, type = "normalized"))
```


```{r}
m2 <- gamm(Annual ~ s(Year, k = 30), data = gtemp,
           correlation = corARMA(form = ~ Year, p = 1))
## ...and fit the AR2
m3 <- gamm(Annual ~ s(Year, k = 30), data = gtemp,
           correlation = corARMA(form = ~ Year, p = 2))

 anova(m1$lme, m2$lme, m3$lme)
plot(m2$gam, residuals = TRUE, pch = 19, cex = 0.75)
with(gtemp, tsDiagGamm(m2, timevar = Year, observed = Annual))
```


The fitted model may be overly complex and over fitted to the data if we do not account for the violation of independence in the residuals.

```{r}
plot(Annual ~ Year, data = gtemp, type = "p", ylab = ylab)
pdat <- with(gtemp,
             data.frame(Year = seq(min(Year), max(Year),
                        length = 200)))
p1 <- predict(m1$gam, newdata = pdat)
p2 <- predict(m2$gam, newdata = pdat)
lines(p1 ~ Year, data = pdat, col = "red")
lines(p2 ~ Year, data = pdat, col = "blue")
legend("topleft",
       legend = c("Uncorrelated Errors","AR(1) Errors"),
       bty = "n", col = c("red","blue"), lty = 1)
```


when were temperatures statistically significantly increasing

method of finite differences

To produce derivatives via finite differences, we compute the values of the fitted trend at a grid of points over the entire data. We then shift the grid by a tiny amount and recompute the values of the trend at the new locations.  The differences between the two sets of fitted values are the first differences of the trend and give a measure of the slope of the trend at any point in time.

```{r}
MyDeriv <- function(mod, n = 200, eps = 1e-7, newdata) {
    if(isTRUE(all.equal(class(mod), c("gamm", "list"))))
        mod <- mod$gam
    m.terms <- attr(terms(mod), "term.labels")
    if(missing(newdata)) {
        newD <- sapply(model.frame(mod)[, m.terms, drop = FALSE],
                       function(x) seq(min(x), max(x), length = n))
        names(newD) <- m.terms
    } else {
        newD <- newdata
    }
    X0 <- predict(mod, data.frame(newD), type = "lpmatrix")
    newD <- newD + eps
    X1 <- predict(mod, data.frame(newD), type = "lpmatrix")
    Xp <- (X1 - X0) / eps
    Xp.r <- NROW(Xp)
    Xp.c <- NCOL(Xp)
    ## dims of bs
    bs.dims <- sapply(mod$smooth, "[[", "bs.dim") - 1
    # number of smooth terms
    t.labs <- attr(mod$terms, "term.labels")
    nt <- length(t.labs)
    ## list to hold the derivatives
    lD <- vector(mode = "list", length = nt)
    names(lD) <- t.labs
    for(i in seq_len(nt)) {
        Xi <- Xp * 0
        want <- grep(t.labs[i], colnames(X1))
        Xi[, want] <- Xp[, want]
        df <- Xi %*% coef(mod)
        df.sd <- rowSums(Xi %*% mod$Vp * Xi)^.5
        lD[[i]] <- list(deriv = df, se.deriv = df.sd)
        ## Xi <- Xp * 0 ##matrix(0, nrow = Xp.r, ncol = Xp.c)
        ## J <- bs.dims[i]
        ## Xi[,(i-1) * J + 1:J + 1] <- Xp[,(i-1) * J + 1:J +1]
        ## df <- Xi %*% coef(mod)
        ## df.sd <- rowSums(Xi %*% mod$Vp * Xi)^.5
        ## lD[[i]] <- list(deriv = df, se.deriv = df.sd)
    }
    class(lD) <- "Deriv"
    lD$gamModel <- mod
    lD$eps <- eps
    lD$eval <- newD - eps
    return(lD)
}
```

```{r}
m2.d <- MyDeriv(m2, n = 200)
plot(m2.d, sizer = TRUE, alpha = 0.01)
```







```{r}
plot(Annual ~ Year, data = gtemp, type = "p", ylab = ylab)
lines(p2 ~ Year, data = pdat)
CI <- confint(m2.d, alpha = 0.01)
S <- signifD(p2, m2.d$Year$deriv, CI$Year$upper, CI$Year$lower,
             eval = 0)
lines(S$incr ~ Year, data = pdat, lwd = 3, col = "blue")
lines(S$decr ~ Year, data = pdat, lwd = 3, col = "red")
```




