---
title: "Data wrangling"
author: "Mark Jones"
date: "`r Sys.time()`"
output:
  html_document:
    classoption: landscape
    css: style.css
    number_sections: yes
    self_contained: yes
    theme: united
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document:
    toc: yes
    toc_depth: '3'
geometry: left=0.2cm,right=0.2cm,top=1cm,bottom=1cm
editor_options:
  chunk_output_type: console
classoption: landscape
---

<!--    toc: yes
    toc_float: true -->

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = 'figs/')
# install.packages(libname, dependencies = T, repos = "http://cran.curtin.edu.au")
suppressPackageStartupMessages(library(simstudy))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(mcmc))
suppressPackageStartupMessages(library(brms))
suppressPackageStartupMessages(library(forecast))
suppressPackageStartupMessages(library(ggfortify))
suppressPackageStartupMessages(library(readxl))


ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text = element_text(size = 10))
ggplot2::theme_update(legend.position = "top")
# ggplot2::theme_update(legend.title = element_blank())
ggplot2::theme_update(axis.text.x = element_text(size = 10))
ggplot2::theme_update(axis.text.y = element_text(size = 10))

# Work|Right|Fast
# rmarkdown::render("simulation_report.Rmd", clean=TRUE)
cbp <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")


```

# Introduction

Consolidated set of data manipulations to generate the analysis dataset.

+ Add indicator to exclude first year of vaccine from our analysis. We look at effects from Jul 2008.
+ Seasonal indicators added (Winter, Spring, Summer, Autumn)
+ Outbreak appears to be restricted to Jul - Oct 2010 inclusive in "3 years", "4 years" and "5-9 years". Outbreak indicators applied to these months/agegrps.


```{r, echo = T}
d <- readxl::read_xls("dat_orig/Times series set up_2019-04-14.xls")
# by agegrp
d1 <- d %>% 
  # the last count had incomplete data
  dplyr::filter(!(year_adm == 2012 & month_adm == 12)) %>%
  # i_vac is an indicator of the vaccine year
  dplyr::rename(i_vac = vaccineyear,
                count = counts,
                mnth = month_adm,
                yr = year_adm) %>%
  dplyr::mutate(agegrp = factor(agegpfull, 
                                labels = unique(d$agegpfull),
                                levels = unique(d$agegpfull))) %>%
  dplyr::select(-agegpfull, 
                -denom_nonindig,
                -denom_indig,
                -counts_nonindig,
                -counts_indig,
                -monthcount) %>%
  dplyr::arrange(agegrp, yr, mnth) %>%
  # per 100,000
  dplyr::mutate(rate = 100000 * count / denom) %>%
  dplyr::group_by(agegrp) %>%
  # mnth_sequence is the time point for each observation in each group
  # mnth_since_vac is months from time vaccine introduced
  dplyr::mutate(mnth_seq = 0:(n()-1),
                mnth_since_vac = 0:(n()-1), 
                # effects of vac considered from Jul 2008
                mnth_since_vac = mnth_since_vac -48) %>%
  dplyr::ungroup() %>%
  # get rid of the negative values in mnth_since_vac
  dplyr::mutate(mnth_since_vac = ifelse(mnth_since_vac < 0, 0, mnth_since_vac),
                incl = ifelse(mnth_seq %in% 36:47, 0, 1)) %>%
  dplyr::group_by(agegrp) %>%
  # add some additional age group level variates
  dplyr::mutate(rate_lag1 = lag(rate),
                rate_d1 = rate - lag(rate),
                rate_d1_lag1 = lag(rate_d1)) %>%
  dplyr::ungroup()

# add seaonal indicators
d1 <- d1 %>%
  dplyr::mutate(winter = ifelse(mnth %in% c(6, 7, 8), 1, 0),
                spring = ifelse(mnth %in% c(9, 10, 11), 1, 0),
                summer = ifelse(mnth %in% c(12, 1, 2), 1, 0)) %>%
  dplyr::mutate(season = ifelse(winter == 1, "Win", 
                                ifelse(spring == 1, "Spr", 
                                       ifelse(summer == 1, "Sum", "Aut"))))
d1$season <- factor(d1$season, levels = c("Aut", "Win", "Spr", "Sum"))

# add harmonic - not used but retained anyway
d1$cos.t1 <- cos(2*pi*d1$mnth_seq*(10/120))
d1$sin.t1 <- sin(2*pi*d1$mnth_seq*(10/120))
d1$cos.t2 <- cos(2*pi*d1$mnth_seq*(20/120))
d1$sin.t2 <- sin(2*pi*d1$mnth_seq*(20/120))

d1$outbreak <- 0

d1$outbreak[d1$yr == 2010 & 
              d1$agegrp %in% c("3 years", "4 years", "5-9 years")] <- 1


saveRDS(d1, "dat_clean\\dat_2019-04-14.RDS")

```




