---
title: "EDA"
author: "Mark Jones"
date: "`r Sys.time()`"
output:
  html_document:
    classoption: landscape
    css: style.css
    number_sections: yes
    self_contained: yes
    theme: united
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document:
    toc: yes
    toc_depth: '3'
geometry: left=0.2cm,right=0.2cm,top=1cm,bottom=1cm
editor_options:
  chunk_output_type: console
classoption: landscape
---

<!--    toc: yes
    toc_float: true -->

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = 'figs/')
# install.packages(libname, dependencies = T, repos = "http://cran.curtin.edu.au")
suppressPackageStartupMessages(library(simstudy))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(mcmc))
suppressPackageStartupMessages(library(brms))
suppressPackageStartupMessages(library(forecast))

ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text = element_text(size = 10))
ggplot2::theme_update(legend.position = "top")
# ggplot2::theme_update(legend.title = element_blank())
ggplot2::theme_update(axis.text.x = element_text(size = 10))
ggplot2::theme_update(axis.text.y = element_text(size = 10))

# Work|Right|Fast
# rmarkdown::render("simulation_report.Rmd", clean=TRUE)
cbp <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")


```


# Preamble

Load data, filter out last month, compute rate (per 100k), introduce month sequence number (starts at 0) and time since introduction of RV1 (starts at zero). Aggregate across all age-groups.

```{r, echo = T}
d <- readxl::read_xls("dat_orig\\Times series set up.xls")
# by agegrp
d1 <- d %>% 
  dplyr::filter(!(year_adm == 2012 & month_adm == 12)) %>%
  dplyr::rename(i_vac = vaccineyear,
                count = counts,
                mnth = month_adm,
                yr = year_adm) %>%
  dplyr::mutate(agegrp = factor(agegpfull, 
                                labels = unique(d$agegpfull),
                                levels = unique(d$agegpfull))) %>%
  dplyr::select(-agegpfull, 
                -denom_nonindig,
                -denom_indig,
                -counts_nonindig,
                -counts_indig,
                -monthcount) %>%
  dplyr::arrange(agegrp, yr, mnth) %>%
  dplyr::mutate(rate = 100000 * count / denom) %>%
  dplyr::group_by(agegrp) %>%
  dplyr::mutate(mnth_seq = 0:(n()-1),
                mnth_since_vac = 0:(n()-1), 
                mnth_since_vac = mnth_since_vac -36) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(mnth_since_vac = ifelse(i_vac == 0, 0, mnth_since_vac),
                rate_lag1 = lag(rate),
                rate_d1 = rate - lag(rate),
                rate_d1_lag1 = lag(rate_d1)) 

# aggregated across all agegrp
d2 <- d1 %>% 
  dplyr::group_by(yr, mnth, i_vac) %>%
  dplyr::summarise(count = sum(count),
                   denom = sum(denom)) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(yr, mnth) %>%
  dplyr::mutate(rate = 100000 * count / denom,
                mnth_seq = 0:(n()-1),
                mnth_since_vac = 0:(n()-1)) %>%
  dplyr::mutate(mnth_since_vac = ifelse(i_vac == 0, 0, mnth_since_vac),
                rate_lag1 = lag(rate),
                rate_d1 = rate - lag(rate),
                rate_d1_lag1 = lag(rate_d1)) 

saveRDS(d1, "dat_clean\\d_by_age.RDS")
saveRDS(d2, "dat_clean\\d_all_age.RDS")
```


# Visualisation

Gastro presentations by age-group. 

Non-stationary in level and variance.

```{r, echo = F, fig.width=10, fig.height=9}
ggplot(d1, aes(x = mnth_seq, y = rate))+
  geom_line()+
  geom_vline(xintercept = 37, colour = cbp[2]) +
  scale_x_continuous("Months from Jul'04") +
  scale_y_continuous("Rate per 100k")+
  facet_wrap(~agegrp)
```

Allow different y-axis for each age group. What is the peak attributable to that is around month 75 in the 3 year, 4 year and 5-9 year age group? 

```{r, echo = F, fig.width=10, fig.height=9}
ggplot(d1, aes(x = mnth_seq, y = rate))+
  geom_line()+
  geom_vline(xintercept = 37, colour = cbp[2]) +
  scale_x_continuous("Months from Jul'04") +
  scale_y_continuous("Rate per 100k")+
  facet_wrap(~agegrp, scales = "free")
```

All-age gastro presentations.

```{r, echo = F, fig.width=7, fig.height=4}
ggplot(d2, aes(x = mnth_seq, y = rate))+
  geom_line()+
  geom_vline(xintercept = 37, colour = cbp[2]) +
  scale_x_continuous("Months from Jul'04") +
  scale_y_continuous("Rate per 100k")
```

Monthly seasonality not readily apparent. Maybe some increase towards the end of the year in younger age groups.

```{r, echo = F, fig.width=10, fig.height=9}
ggplot(d1, aes(x = factor(mnth), y = rate, group = yr, colour = factor(yr)))+
  geom_line()+
  scale_x_discrete("Months", breaks = seq(0, 12, 2)) +
  scale_y_continuous("Rate per 100k")+
  scale_color_discrete("Year")+
  facet_wrap(~agegrp, scales = "free")
```

First differences.

```{r, echo = F, fig.width=10, fig.height=9}
ggplot(d1 %>% 
         dplyr::group_by(agegrp) %>%
         dplyr::filter(row_number() != 1) %>%
         dplyr::ungroup(), 
       aes(x = mnth_seq, y = rate_d1))+
  geom_line()+
  geom_vline(xintercept = 37, colour = cbp[2]) +
  scale_x_continuous("Months from Aug'04") +
  scale_y_continuous("Rate per 100k")+
  facet_wrap(~agegrp, scales = "free")
```

ACFs.

```{r, echo = F, fig.width=10, fig.height=9}
# https://stackoverflow.com/questions/37325517/acf-by-group-in-r
# https://rh8liuqy.github.io/ACF_PACF_by_ggplot2.html
d_acf <- d1 %>%
  dplyr::group_by(agegrp) %>%
  dplyr::filter(row_number() != 1) %>%
  tidyr::nest(-agegrp) %>%
  dplyr::mutate(acf_results = purrr::map(data, ~ acf(.x$rate_d1, plot = F)),
                acf = purrr::map(acf_results, ~ drop(.x$acf))) %>%
  tidyr::unnest(acf) %>%
  dplyr::group_by(agegrp) %>%
  dplyr::mutate(lag = seq(0, n() - 1))

ggplot(data = d_acf, 
            aes(x = lag, y = acf)) +
  geom_hline(aes(yintercept = 0)) +
  scale_x_continuous("Months lag") +
  geom_segment(mapping = aes(xend = lag, yend = 0)) + 
  facet_wrap(~agegrp)

```

Lag-1 plots (non-stationary level) - how the present rate relates to the previous month's rate in the rate series. Mainly reflects the trend.

```{r, echo = F, fig.width=10, fig.height=9}
# http://rpubs.com/sinhrks/plot_tsstats
ggplot(d1 %>% 
         dplyr::group_by(agegrp) %>%
         dplyr::filter(row_number() != 1) %>%
         dplyr::ungroup(), 
       aes(x = rate_lag1, y = rate))+
  geom_point()+
  geom_smooth(method = "lm")+
  scale_x_continuous("Rate @ lag 1") +
  scale_y_continuous("Rate")+
  facet_wrap(~agegrp, scales = "free")

```

Lag-1 plots (stationary level)- how the present rate relates to the previous month's rate in the first differences. Some negative autocorrelation in older age groups (high rate one month associated with low rates the next month). Not much (that I really have faith in) apparent. 

```{r, echo = F, fig.width=10, fig.height=9}
# http://rpubs.com/sinhrks/plot_tsstats
ggplot(d1 %>% 
         dplyr::group_by(agegrp) %>%
         dplyr::filter(row_number() != 1 & row_number() != 2) %>%
         dplyr::ungroup(), 
       aes(x = rate_d1_lag1, y = rate_d1))+
  geom_point()+
  geom_smooth(method = "lm")+
  scale_x_continuous("Rate @ lag 1") +
  scale_y_continuous("Rate")+
  facet_wrap(~agegrp, scales = "free")

```


