---
title: "GLM - seasonality"
author: "Mark Jones"
date: "`r Sys.time()`"
output:
  html_document:
    classoption: landscape
    css: style.css
    number_sections: yes
    self_contained: yes
    theme: united
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document:
    toc: yes
    toc_depth: '3'
geometry: left=0.2cm,right=0.2cm,top=1cm,bottom=1cm
editor_options:
  chunk_output_type: console
classoption: landscape
---

<!--    toc: yes
    toc_float: true -->

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = 'figs/')
# install.packages(libname, dependencies = T, repos = "http://cran.curtin.edu.au")
suppressPackageStartupMessages(library(simstudy))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(mcmc))
suppressPackageStartupMessages(library(brms))
suppressPackageStartupMessages(library(forecast))
suppressPackageStartupMessages(library(ggfortify))
suppressPackageStartupMessages(library(readxl))


ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text = element_text(size = 10))
ggplot2::theme_update(legend.position = "top")
# ggplot2::theme_update(legend.title = element_blank())
ggplot2::theme_update(axis.text.x = element_text(size = 10))
ggplot2::theme_update(axis.text.y = element_text(size = 10))

# Work|Right|Fast
# rmarkdown::render("simulation_report.Rmd", clean=TRUE)
cbp <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")


```


# Preamble

Several approaches are available to model seasonality ranging from deterministic indicators to stochastic autoregressive components.

We use the data that is configured to exclude the first year of vax.

## Seasonality

For the first go at introducing seasonality, add indicator variables to represent seasons to the model. December to February is summer; March to May is autumn; June to August is winter; and September to November is spring. The benefit with this approach is that it is very easily interpretable. The downside is that it probably doesn't model seasonality optimally.

In addition to the exisiting covariates, this new specification allows differential pre and post seasonal effects. However, within the pre and post period the seasonal effects are fixed (they cannot vary over time).

```{r, echo = T, eval = T}
d1 <- readRDS("dat_clean//d_rm_vax_1.RDS")

# Autumn corresponds to the reference season
d1 <- d1 %>%
  dplyr::mutate(winter = ifelse(mnth %in% c(6, 7, 8), 1, 0),
                spring = ifelse(mnth %in% c(9, 10, 11), 1, 0),
                summer = ifelse(mnth %in% c(12, 1, 2), 1, 0)) %>%
  dplyr::mutate(season = ifelse(winter == 1, "Win", 
                                ifelse(spring == 1, "Spr", 
                                       ifelse(summer == 1, "Sum", "Aut"))))
d1$season <- factor(d1$season, levels = c("Aut", "Win", "Spr", "Sum"))

saveRDS(d1, "dat_clean\\d_seas1.RDS")
```

# GLM

Independent poisson models fitted for each age group. Exclude the first year of vax from the model and include seasonal indicator covariates.

```{r}
fit1 <- function(x){
  dfit <- d1 %>%
    dplyr::filter(agegrp == x & incl == 1)
  lm1 <- glm(count ~ mnth_seq + factor(i_vac) + mnth_since_vac + 
               winter + spring + summer +
               offset(log(denom)), 
             data = dfit, family = "poisson")
  
  # in order to get the rate (number of cases of one unit of population)
  # you need to do this hack
  dpred <- d1 %>%
    dplyr::filter(agegrp == x)  %>%
    dplyr::mutate(denom = 1)
  
  d <- d1 %>%
    dplyr::filter(agegrp == x)
  
  d$pred <- predict(lm1, type = "response", newdata = dpred) * 100000
  d$pred <- ifelse(d$mnth_seq %in% 36:47, NA, d$pred)
  
  list(desc = paste0(x), lm1, d)
}
fits <- lapply(unique(d1$agegrp), fit1)

```

Model results. Interpretation is described previously. AIC included.

```{r}
get_dat <- function(x){
  return(list(desc = x[[1]], 
              model_res = broom::tidy(x[[2]]),
              aic = AIC(x[[2]])))
}
l_res <-lapply(fits, get_dat)
l_res
```

## Plot predictions.

Again, we have a gap for the Jul 2007 to Jun 2008 period. The seasonal effects are shown. Winter and Spring appear to consistently be the seasons in which most cases occur. It looks like the pre-vax Winter and Spring peaks have been damped down in the post vax period. This is most apparent in the lower ages.

```{r, echo = F, fig.width=10, fig.height=9}
get_dat <- function(x){
  return(x[[3]])
}
d_res <- dplyr::bind_rows(lapply(fits, get_dat))
d_res$season <- factor(d1$season, 
                       levels = c("Aut", "Win", "Spr", "Sum"),
                       labels = c("Aut", "Win", "Spr", "Sum"))

ggplot(d_res, aes(x = mnth_seq, y = rate))+
  geom_line()+
  geom_line(aes(x = mnth_seq, y = pred, colour = cbp[2]), size = 1) + 
  geom_point(aes(x = mnth_seq, y = pred, colour = season)) +
  scale_color_discrete(breaks=c("Aut", "Win", "Spr", "Sum"))+
  geom_vline(xintercept = 37, colour = cbp[2]) +
 # theme(legend.position = "none") +
  scale_x_continuous("Months from Jul'04") +
  scale_y_continuous("Rate per 100k")+
  facet_wrap(~agegrp, scales = "free", ncol = 3)
```

# Summary

Simple seasonal indicator variables included.


