---
title: "GLARMA - poisson"
author: "Mark Jones"
date: "`r Sys.time()`"
output:
  html_document:
    classoption: landscape
    css: style.css
    number_sections: yes
    self_contained: yes
    theme: united
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document:
    toc: yes
    toc_depth: '3'
geometry: left=0.2cm,right=0.2cm,top=1cm,bottom=1cm
editor_options:
  chunk_output_type: console
classoption: landscape
---

<!--    toc: yes
    toc_float: true -->

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = 'figs/')
# install.packages(libname, dependencies = T, repos = "http://cran.curtin.edu.au")
suppressPackageStartupMessages(library(simstudy))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(mcmc))
suppressPackageStartupMessages(library(brms))
suppressPackageStartupMessages(library(forecast))
suppressPackageStartupMessages(library(ggfortify))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(glarma))

ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text = element_text(size = 10))
ggplot2::theme_update(legend.position = "top")
# ggplot2::theme_update(legend.title = element_blank())
ggplot2::theme_update(axis.text.x = element_text(size = 10))
ggplot2::theme_update(axis.text.y = element_text(size = 10))

# Work|Right|Fast
# rmarkdown::render("simulation_report.Rmd", clean=TRUE)
cbp <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")


```


# Preamble

GLARMA (Generalised Linear Auto-regressive Moving average) models are a more appropriate way to include AR and MA components (better than just using a lag-1 variable as previously shown).

http://www.stat.columbia.edu/~rdavis/lectures/ambleside00.pdf


```{r, echo = T, eval = T}
d1 <- readRDS("dat_clean//d_seas2.RDS")
d1$rate_lag1 <- d1$rate_lag1  / 1000
```

# GLARMA Models

Independent poisson models fitted for each age group. To get these models to fit you need to produce a GLM to get rough estimates of the parameters and then use these estimates as the initial values for betas in the GLARMA models.

The results suggest that incorporating the AR1 term is generally justified.

```{r}
fit1 <- function(x){
  dfit <- d1 %>%
    dplyr::filter(agegrp == x & incl == 1)

  lm1 <- glm(count ~ mnth_seq + i_vac + mnth_since_vac + 
               winter + spring + summer +  
               offset(log(denom)), 
             data = dfit, family = "poisson")
  
  mX <- model.matrix(lm1)
  y <- dfit$count
  mOffset <- log(dfit$denom)
  
  lm2 <- glarma(y, mX, offset = mOffset,
              phiLags = c(1),
              #thetaLags = 3, #thetaInit = c(0.01),
              type = "Poi",
              # alpha = glm1$theta, alphaInit = glm1$theta,
              beta = coef(lm1),
              method = "FS",
              residuals = "Pearson", maxit = 1000, grad = 2.22e-16)
  s <- summary(lm2)
  
  # fitted
  dpred <- d1 %>%
    dplyr::filter(agegrp == x)  %>%
    dplyr::mutate(denom = 1)
  
  d <- d1 %>%
    dplyr::filter(agegrp == x)
  
  d$pred1 <- predict(lm1, type = "response", newdata = dpred) * 100000
  d$pred1 <- ifelse(d$mnth_seq %in% 36:47, NA, d$pred1)
  
  pred2 <- 100000 * glarma::fitted.glarma(lm2)/dfit$denom
  
  d$pred2 <- NA
  d$pred2[1:36] <- pred2[1:36]
  d$pred2[49:nrow(d)] <- pred2[37:length(pred2)]
  
  list(desc = paste0(x), lm1, lm2, d)
}
fits <- lapply(unique(d1$agegrp), fit1)

```

Model results. Interpretation is described previously.

The AIC values are consistently better (lower) than the previous models that just use the indicator variables.

```{r, eval = T}
get_dat <- function(x){
  return(list(desc = x[[1]], 
              glm_res = summary(x[[2]]),
              glarma_res = summary(x[[3]])))
}
l_res <-lapply(fits, get_dat)
l_res
```

## Plot predictions.

Again, we have a gap for the Jul 2007 to Jun 2008 period. The seasonal effects are shown. Winter and Spring appear to consistently be the seasons in which most cases occur. It looks like the pre-vax Winter and Spring peaks have been damped down in the post vax period. This is most apparent in the lower ages.

```{r, echo = F, fig.width=9, fig.height=12, eval = T}
get_dat <- function(x){
  return(x[[4]])
}
d_res <- dplyr::bind_rows(lapply(fits, get_dat))


ggplot(d_res, aes(x = mnth_seq, y = rate))+
  geom_line()+
  geom_line(aes(x = mnth_seq, y = pred1, colour = cbp[2]), size = 0.8) + 
  geom_line(aes(x = mnth_seq, y = pred2, colour = cbp[6]), size = 0.8) + 
  geom_vline(xintercept = 37, colour = cbp[2]) +
 # theme(legend.position = "none") +
  scale_x_continuous("Months from Jul'04") +
  scale_y_continuous("Rate per 100k")+
  facet_wrap(~agegrp, scales = "free", ncol = 2)
```

## Diagnostics

These show that poisson distributional assumption is not really appropriate for this data.

```{r, eval = T}

for(i in 1:length(fits)){
  l1 <- fits[[i]][[3]]
  par(mfrow = c(2,2))
  plot(l1, which = c(2,3,5),
       titles = list(NULL, 
                     NULL, 
                     "PIT for GLARMA (Poisson)"))
}

```


# Summary

GLARMA models fitted under poisson assumptions. They don't fit very well. 


