---
title: "BSTS"
author: "Mark Jones"
date: "`r Sys.time()`"
output:
  html_document:
    classoption: landscape
    css: style.css
    number_sections: yes
    self_contained: yes
    theme: united
  pdf_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document:
    toc: yes
    toc_depth: '3'
geometry: left=0.2cm,right=0.2cm,top=1cm,bottom=1cm
editor_options:
  chunk_output_type: console
classoption: landscape
---

<!--    toc: yes
    toc_float: true -->

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = 'figs/')
# install.packages(libname, dependencies = T, repos = "http://cran.curtin.edu.au")
suppressPackageStartupMessages(library(simstudy))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(mcmc))
suppressPackageStartupMessages(library(brms))
suppressPackageStartupMessages(library(forecast))
suppressPackageStartupMessages(library(ggfortify))
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(geepack))

ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text = element_text(size = 10))
ggplot2::theme_update(legend.position = "top")
# ggplot2::theme_update(legend.title = element_blank())
ggplot2::theme_update(axis.text.x = element_text(size = 10))
ggplot2::theme_update(axis.text.y = element_text(size = 10))

# Work|Right|Fast
# rmarkdown::render("simulation_report.Rmd", clean=TRUE)
cbp <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")


```


# Preamble

Bayesian methods include structural time series models.

```{r, echo = T, eval = T}
d1 <- readRDS("dat_clean//d_seas2.RDS")
d1$rate_lag1 <- d1$rate_lag1  / 1000
```

# BSTS

I think these are not yet fully supported in `brms`. Just stick with `glarma` for now.

On testing the model I get the following warning:

> 1: The 'bsts' correlation structure has been deprecated and will be removed from the package at some point. Consider using splines or Gaussian processes instead.

So splines and gaussian processes may be worth a look.

```{r, eval = F}
fit1 <- function(x){
  dfit <- d1 %>%
    dplyr::filter(agegrp == x & incl == 1)

  lm1 <- brm(count ~ mnth_seq + i_vac + mnth_since_vac + 
               winter + spring + summer + 
               offset(log(denom)), 
             data = dfit, autocor = cor_bsts(formula = ~ mnth_seq),
             family = "negbinomial")
      
  
  list(desc = paste0(x), lm1)
}
fits <- lapply(unique(d1$agegrp), fit1)

```

# Summary

I like the idea of taking a Bayesian approach to this but that is probably purely for my own interest rather than necessary.




